---
TASK: '2'
Author: Shreya Khare
---
# SHREYA KHARE

# TASK 2

#Importing Libraries
```{r}
library(vegan)
library(permute)
library(lattice)
library(NbClust)
library(factoextra)
library(stats19)
library(dplyr)
library(ggplot2)
library(ggfortify)
```


#loading datasset
```{r}
df <- read.csv("Iris.csv")
head(df)
```


#structure of data
```{r}
str(df)
```


#summary of dataset
```{r}
summary(df)
```



```{r}
dt<- select(df,c(1,2,3,4))
```


#Scatter plot
```{r}
scatter <- ggplot(data=iris, aes(x = Petal.Length, y = Petal.Width)) 
res <- scatter + geom_point(aes(color=Species, shape=Species)) +
  theme_bw()+
  xlab("Petal Length") +  ylab("Petal Width") +
  ggtitle("Petal Length-Width")+
  theme(plot.title = element_text(hjust = 0.5))
res
```


#Within group Sum of Squares (WSS) plot to determine the optimal no of clusters
```{r}
wss <- (nrow(dt)-1)*sum(apply(dt,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(dt,
                                     centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares", col = "red")

```


#NbClust funtion for determining the best number of clusters
```{r}
par(mar = c(2,2,2,2))
irisData <- iris[,1:4]
totalwSS<-c()
nb <- NbClust(irisData, method = "kmeans")
```


#histogram approach denoting various indices with different number of clusters voted
```{r}
hist(nb$Best.nc[1,], breaks = 15, main="Number of Clusters",col = c("blue", "red", "gold", "green")) 

```


#Calinski-Harabasz index the Variance Ratio Criterion

#It is the ratio of inter-cluster dispersion for all clusters

#Higher the score , the better the performances
```{r}
modelData <- cascadeKM(irisData, 1, 15, iter = 100)
plot(modelData, sortg = TRUE)
modelData$results[2,]

```


#To determine which of these values is maximum? BC/WC should be as large as possible
```{r}
which.max(modelData$results[2,])
```